Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-03-30 ThesisJP/Background/background.tex
Added new figure to the medical ulttrasound imaging sector.

\section {Medical ultrasound imaging}\label{sec:ultrasound}
Ultrasound imaging encompasses technology which generates images based on sound whose frequencies we can not hear. For medical ultrasound imaging frequencies between 2 MHz and 10 MHz are typically used. Which frequency to select depends on how deep into the body we want to image, this because human tissue exhibits frequency dependent attenuation. Frequency dependent attenuation is the reason why a frequency around 3 MHz is used for fetal and cardiac imaging and a frequency around 7 MHz is used for more shallow imaging (e.g. vascular imaging).

Medical ultrasound imaging is considered to be a low-risk tool for non-invasive investigation. There are no gamma rays or high intensity magnetic fields involved, just sound pressure below well controlled limits. It is also far less costly than other imaging modalities and offers real-time inspection of the human body.  Ultrasound imaging is well know for its use in fetal imaging during pregnancy, however, in this thesis we will focus on a different modality, namely cardiac ultrasound imaging. An illustration of the human heart is therefore included in Fig.\,\ref{fig:human_heart}. Throughout the thesis we will refer to some of the parts shown in Fig.\,\ref{fig:human_heart}. Especially note the four chambers of the heart, the left and right atrium, and the left and right ventricle. The tissue separating the left and right ventricle is know as the inter-ventricular septum. The left side is the strongest and hardest working part of the heart, and it is responsible of pumping blood out to the whole body through the aortic valve. When blood flows out the aorta it is important that the mitral valve, that separates the left atrium and ventricle, is closed and does not leak. The left ventricle and the mitral valve therefore receives special attention during cardiac ultrasound examinations. When visualizing 3D ultrasound images in \textbf{Paper IV} we will have a strong focus on visibility of cardiac tissue, thus note for later that the endocardium refers to the innermost layer of the cardiac muscle.

\begin{figure}
\centering
\subfigure{
	\includegraphics[width=0.47\textwidth]{img/Diagram_of_the_human_heart.png}
}
\subfigure{
	\includegraphics[width=0.47\textwidth]{img/HeartWall.png}
}
\caption{Overview of the human heart. Illustrations from wikipedia.org.}
\label{fig:human_heart}
\end{figure}

To generate the ultrasonic frequencies used for imaging a piezoelectric ceramic is used. A piezoelectric material is special in that way that ultrasound vibrations are generated when voltage is applied and \textit{vice versa}. It can therefore be used to both send and receive ultrasound signals, like a combined microphone and speaker. Piezoelectric elements organized in an array, is referred to as an ultrasound \textit{probe}, and a special probe is typically designed for each medical imaging modality. For cardiac ultrasound imaging the probe must not only be designed for a certain frequency range, its footprint also needs to be smaller than the intercostal window in order to avoid shadowing by ribs. A line in an ultrasound image is generated by firing a ultrasound pulse into the body and sample the received echo. This amplitude trace is known as an A-mode image.

>>>>

\section {Medical ultrasound imaging}\label{sec:ultrasound}
Medical ultrasound imaging is considered to be a low-risk tool for non-invasive investigation. Neither gamma rays nor high intensity magnetic fields are involved, just sound pressure below well-controlled limits. It is also far less costly than other medical imaging modalities and offers real-time inspection of the human body.  Ultrasound imaging is well known for its use in fetal imaging during pregnancy, however, in this thesis we will focus on a different modality, namely cardiac ultrasound imaging. An illustration of the human heart is therefore included in Fig.\,\ref{fig:human_heart}. Throughout the thesis we will refer to some of the parts shown in Fig.\,\ref{fig:human_heart}. Especially note the four chambers of the heart, the left and right atrium, and the left and right ventricle. The tissue separating the left and right ventricle is know as the inter-ventricular septum. The left side is responsible of pumping blood out to the whole body through the aortic valve. It is therefore the strongest and hardest working part of the heart. When blood flows out of the aorta it is important that the mitral valve, which separates the left atrium from the left ventricle, is closed and does not leak. The left side, and especially the left ventricle and the mitral valve, therefore receives the most attention during cardiac ultrasound examinations. When visualizing 3D ultrasound images in \textbf{Paper IV} we will have a strong focus on visibility of cardiac tissue, thus note for later that the endocardium refers to the innermost layer of the cardiac muscle.

\begin{figure}
\centering
\subfigure[]{
	\includegraphics[width=0.47\textwidth]{img/Diagram_of_the_human_heart.png}\label{fig:human_heart}
}
\subfigure[]{
	\includegraphics[width=0.47\textwidth]{img/ultrasound_grid.eps}\label{fig:us_grid}
}
\caption{a) Overview of the human heart. Illustration from wikipedia.org. b) Coordinate system for phased array imaging.}
\end{figure}

%Ultrasound imaging encompasses technology which generates images based on sound whose frequencies we can not hear. For medical ultrasound imaging frequencies between 2 MHz and 10 MHz are typically used. The choice of frequency depends on the depth of the body layers we want to image, because human tissue exhibits frequency dependent attenuation. This is why a frequency around 3 MHz is used for fetal and cardiac imaging and a frequency around 7 MHz is used for more shallow imaging (e.g. vascular imaging).

To generate the ultrasonic frequencies used for imaging a piezoelectric ceramic is used. A piezoelectric material is special in that way that ultrasound vibrations are generated when voltage is applied and \textit{vice versa}. It can therefore be used to both send and receive ultrasound signals, like a combined microphone and speaker. Piezoelectric elements organized in an array, is referred to as an ultrasound \textit{probe}, and a special probe is typically designed for each medical imaging modality. %For cardiac ultrasound imaging the probe must not only be designed for a certain frequency range, its footprint also needs to be smaller than the intercostal window in order to avoid shadowing by ribs. 
A line in an ultrasound image is generated by firing a ultrasound pulse into the body and sample the received echo. This amplitude trace is known as an A-mode image.
